Tree height exploration

### 2021-06-02 R/DraftFunctions.Rmd
loggingparameters and addtreedim functions

# My custom options
faut ptetre mettre Maria. devant  chaque objet!
```{r}
# Mimimum DBH value = 10 cm
# Maximum area slope = 27%
# Maximum main&2nd trail centerline slope = 22%
# Maximum main&2nd trail cross slope = 4%
# Grapple maximum slope = 20%
# Trees slope = 22%
# Plateau maximum slope = 5%
# Distance to compute slope = 6 m (3m for each side)
# water sources buffer zone = 30 m
# Main trail width = 5-6 m
# 2nd trail width = 4 m
# Big trees = 50 cm
# Objective bonus = 20-30%
# Cable = 40 m
# Grapple = 6 m
# Isolate tree minimum distance = 100 m
# Future trees minimum diameter = 35 cm
# TreefallSuccess proportion = 0.6
# Treefall orientation = 30-45◦
# (Hollow) Part taken from hollow trees for fuel exploitation = 1/3
# Purge = 0.14
# Maximum TrailDensity = 200 m/ha
# Maximum landing area = 1 500 m^2

# TrunkHeight formule : TrunkHeight = TreeHarvestableVolume/(pi*(((DBH/100)/2)^2))
# CrownDiameter formule : CrownDiameter = exp(((log(DBH)-alpha-error)/beta))/TreeHeight
# RottenModel

pars <- list(MinDBHValue = 10, #in cm
             MaxAreaSlope = 27, #in %
             MaxTrailCenterlineSlope = 22, #in %
             MaxTrailCrossSlope = 4, #in %
             GrappleMaxslope = 20, #in %
             TreeMaxSlope = 22, #in %
             PlateauMaxSlope = 5, #in %
             SlopeDistance = 6, #in m (3m for each side)
             WaterSourcesBufferZone = 30, #in m
             MainTrailWidth = 5-6, #in m
             "2ndTrailWidth" = 4, #in m
             BigTrees = 50, #in cm
             ObjectiveBonus = 20-30, #in % (ObjectiveBonus >= 20 && ObjectiveBonus =< 30 ou c(20:30))
             CableLength = 40, #in m
             GrappleLength = 6, #in m
             IsolateTreeMinDistance = 100, #in m
             FutureTreesMinDiameter = 35, #in cm
             TreefallSuccessProportion = 0.6,
             TreefallOrientation = 30-45, #in degree (◦)
             TreeHollowPartForFuel = 1/3, #(Hollow) Part taken from hollow trees for fuel exploitation
             Purge = 0.14, # in m^3 of fuelwood/m^3 of logged trees  
             MaxTrailDensity = 200, #in m/ha
             MaxLandingArea = 1500, #in square meters
             TrunkHeightAllometry = TreeHarvestableVolume/(pi*(((DBH/100)/2)^2)),
             CrownDiameterAllometry = exp(((log(DBH)-alpha-error)/beta))/TreeHeight,
             RottenModel = ) 
```

@@

@@

#if DBH (cm) doesn't exist create it
if (!("DBH" %in% names(Paracou6_2016))) {add_column(Paracou6_2016, DBH = NA)
  Paracou6_2016$DBH = Paracou6_2016$CircCorr/pi} # and compute it

# TreeHeight = hauteur de l’arbre (H) (en mètre) (BIOMASS package)
# 3 options :
# - get my tree heights from the location (from continent or region specific model, or from coordinates)
# - from other Guyanese trees heights data, fit different H-D models, select the best and apply on my data.

#continent or region specific model (Feldpaush et al. (2012))
RegionTreeHeight <- retrieveH(D = Paracou6_2016$DBH, region = "GuianaShield") # GuianaShield = Guyana, French Guiana (piste St Elie (Chave 2005)), Venezuela

#by coordinates (Chave et al. (2012))
n <- length(Paracou6_2016$CircCor)
coord <- cbind(long = rep(-52.92, n), lat = rep(5.27, n)) # Paracou P6 coordinates
CoordTreeHeight <- retrieveH(D = Paracou6_2016$DBH, coord = coord)

# Just with Nouragues data (measurement: laser rangefinder):
# Fit first different height-diameter models
ModelsresultNouraguesdata <- modelHD( # modelHD = .
  D = NouraguesHD$D,
  H = NouraguesHD$H,
  useWeight = TRUE # model weights will be (D^2)*H (weights are proportional to tree volume, so that larger trees have a stronger influence during the construction of the model)
)
# choose the best model
kable(ModelsresultNouraguesdata) #log2 has the lowest RSE (4.222718) (Root Squared Error)
HDmodelNouraguesdata <- modelHD(
  D = NouraguesHD$D,
  H = NouraguesHD$H,
  method = "log2", # Compute the local H-D model with the lowest RSE.
  useWeight = TRUE
)
# Applicate the model to retrieve tree heights ---- HERE
TreeHeightEstimation <- retrieveH(
  D = Paracou6_2016$DBH,
  model = HDmodelNouraguesdata
)
## Put these tree heights estimations in the inventory
Paracou6_2016TreeHeight <- Paracou6_2016 %>% 
  mutate(TreeHeight = TreeHeightEstimation$H)
plot(Paracou6_2016TreeHeight$DBH, Paracou6_2016TreeHeight$TreeHeight)#amazing !! It's a log2 form

@@

>>>>

@@

```{r}
library(BIOMASS)
library(knitr)

TreeheightGuyafordata <- read_delim("D:/VSC Kourou/DATA/TreeheightGuyafordata.csv", 
                                    ";", escape_double = FALSE, col_types = cols(Circ = col_number(), 
                                                                                 CircCorr = col_number(), Hauteur = col_number()), 
                                    locale = locale(encoding = "Latin1"), 
                                    trim_ws = TRUE) %>% 
  mutate(HeightHow = recode(HeightHow, 'Visual estimate' = 'Visual Estimate')) %>% 
  # filter(!(HeightHow == "Visual Estimate" | HeightHow == "Visual estimate")) %>% 
  unite(Genus, Species, col = "ScientificName", sep = "_", remove = F) %>% 
  mutate(DBH = ifelse(is.na(CircCorr), Circ/pi, CircCorr/pi)) # add DBH column

any(is.na(TreeheightGuyafordata$DBH)) # all the DBH have been computed?

SineTelemeter <- TreeheightGuyafordata %>% #data to fit model
  filter(HeightHow == "Sine Method (Larjavaara & Muller-Landau 2013) with Trimble LaserAce 1000 Rangefinder"| HeightHow == "Telemeter")

ModelsresultGuiana <-modelHD(
  D = SineTelemeter$DBH,
  H = SineTelemeter$Hauteur,
  useWeight = TRUE
)
kable(ModelsresultGuiana) #log2 has the lowest RSE (4.699261 with all data, 4.316927 without "Visual E/estimate", 3.059860 with just "Sine Method", 3.809011 with sine and telemeter)

HDmodelGuiana <- modelHD(
  D = SineTelemeter$DBH,
  H = SineTelemeter$Hauteur,
  method = "log2", # Compute the H-D model with the lowest RSE.
  useWeight = TRUE
)
TreeHeightGuianaEstimation <- retrieveH(
  D = TreeheightGuyafordata$DBH,
  model = HDmodelGuiana
)
#Put these tree heights estimations in the inventory
TreeheightGuyafordataEstim <- TreeheightGuyafordata %>% 
  mutate(HauteurEstimee = TreeHeightGuianaEstimation$H)
plot(TreeheightGuyafordataEstim$DBH, TreeheightGuyafordataEstim$HauteurEstimee)

# obs/estim regression:
TreeheightGuyafordataEstim %>%
  ggplot() +
  aes(x = Hauteur, y = HauteurEstimee) +
  geom_point(shape = "circle", size = 0.5, colour = "#440154") +
  geom_smooth(span = 0.75) + 
  theme_minimal()
# We are well between 10 and 35m height
```

@@

# Just with Nouragues data (measurement: laser rangefinder):
# Fit first different height-diameter models
ModelsresultNouraguesdata <- modelHD(
  D = NouraguesHD$D,
  H = NouraguesHD$H,
  useWeight = TRUE # model weights will be (D^2)*H (weights are proportional to tree volume, so that larger trees have a stronger influence during the construction of the model)
)
# choose the best model
kable(ModelsresultNouraguesdata) #log2 has the lowest RSE (4.222718) (Root Squared Error)
HDmodelNouraguesdata <- modelHD(
  D = NouraguesHD$D,
  H = NouraguesHD$H,
  method = "log2", # Compute the local H-D model with the lowest RSE.
  useWeight = TRUE
)
# Applicate the model to retrieve tree heights ---- HERE
TreeHeightEstimation <- retrieveH(
  D = Paracou6_2016$DBH,
  model = HDmodelNouraguesdata
)
## Put these tree heights estimations in the inventory
Paracou6_2016TreeHeight <- Paracou6_2016 %>% 
  mutate(TreeHeight = TreeHeightEstimation$H)
plot(Paracou6_2016TreeHeight$DBH, Paracou6_2016TreeHeight$TreeHeight)#amazing !! It's a log2 form

@@

  dplyr::rename(Taxo.family = Taxo, alpha.family = alpha, beta.family = beta) %>%
  mutate(alpha = alpha.species, beta = beta.species, Taxo = Taxo.species) %>% # create the future & unique alpha,beta,taxo columns 
  mutate(alpha = ifelse(is.na(alpha), alpha.genus, alpha)) %>% #if species parameters are absent, take genus parameters
  mutate(alpha = ifelse(is.na(alpha), alpha.family, alpha)) %>% #if species&genus parameters are absent, take family parameters
  mutate(alpha = ifelse(is.na(alpha), mean(alpha, na.rm =T), alpha)) %>% #if species,genus&family parameters are absent, take parameters mean
  select(-alpha.species, -alpha.genus, -alpha.family) %>% #remove obsolete columns

### 2021-06-02 R/Tree height exploration.Rmd
loggingparameters and addtreedim functions

any(is.na(TreeheightGuyafordata$DBH)) # all the DBH have been computed?
```

@@

>>>>

any(is.na(TreeheightGuyafordata$DBH)) # all the DBH have been computed?

# DBH range per method
TreeheightGuyafordata %>% 
  group_by(HeightHow) %>% #pour réaliser les opérations suivantes par groupes de modalités
  summarise(MinDBH = min(DBH)) %>% 
  arrange(MinDBH)

TreeheightGuyafordata %>% 
  group_by(HeightHow) %>% #pour réaliser les opérations suivantes par groupes de modalités
  summarise(MaxDBH = max(DBH)) %>% 
  arrange(desc(MaxDBH))
# Visual Estimate: 0.416986 - 256.23946
# Telemeter: 9.867606 - 159.15494
# Optical Telemeter Birnbaum: 7.591691 -131.46198
# Vertex Laser: 10.185916 - 95.73170
# Sine Method: 6.073353 - 88.80846

# measurements number per method
TreeheightGuyafordata %>% 
  count(HeightHow) %>% 
  arrange(desc(n))
# Visual Estimate	: 27986			
# Optical Telemeter Birnbaum: 4902	
# Telemeter: 1454			
# Sine Method:  998	
# Vertex Laser: 441

TreeheightGuyafordata %>% 
  group_by(HeightHow) %>% 
  n_distinct(ScientificName) %>% 
  count(ScientificName) %>% 
  arrange(desc(n))

TreeheightGuyafordata %>% 
  group_by(HeightHow) %>%
  n_distinct(TreeheightGuyafordata$ScientificName) %>% 
  count(ScientificName)
```

@@

```{r}
# Fit Height-Diameter model
library(BIOMASS)

SineTelemeter <- read_delim("D:/VSC Kourou/DATA/TreeheightGuyafordata.csv",
                                    ";", escape_double = FALSE, col_types = cols(Circ = col_number(),
                                                                                 CircCorr = col_number(), Hauteur = col_number()),
                                    locale = locale(encoding = "Latin1"),
                                    trim_ws = TRUE) %>%
  filter(HeightHow == "Sine Method (Larjavaara & Muller-Landau 2013) with Trimble LaserAce 1000 Rangefinder"| HeightHow == "Telemeter") %>%
  mutate(DBH = ifelse(is.na(CircCorr), Circ/pi, CircCorr/pi)) # add DBH column

HDmodel <- modelHD(
  D = SineTelemeter$DBH,
  H = SineTelemeter$Hauteur,
  method = "log2", # Compute the H-D model with the lowest RSE.
  useWeight = TRUE
)

Dmodel[["coefficients"]]

# log 2: log(H) = a + b*log(D) + c*log(D)^2

# log(H) = 0.07359191 + 1.34241216*log(D) + -0.12282344*log(D)^2
# H = exp(0.07359191 + 1.34241216*log(D) + -0.12282344*log(D)^2)

# (Intercept)  0.07359191
# I(log(D)^1)  1.34241216
# I(log(D)^2) -0.12282344
```
